---
title: "Survey 01"
---

## Purpose

This page documents the data processing steps involved with Survey 01 in PSYCH 490.009.

The survey questions were adapted from those discussed in [@Krumrei-Mancuso2016-qm; @Nadelson2014-ym; @Plohl2023-yf].

## Survey

<iframe src="https://docs.google.com/forms/d/e/1FAIpQLScp2GG_2TZ3_T6Yl1okn-sHJF93n9RZu_EmLWQtHP0rJcGM3w/viewform?embedded=true" width="640" height="6003" frameborder="0" marginheight="0" marginwidth="0">Loading…</iframe>

## Preparation

First, we load the external packages (groups of R commands) that we will be using.

```{r include=FALSE}
library('tidyverse')
library('ggplot2')
library('dplyr')
library('tidyr')
library('stringr')
library('lubridate')
```

## Gathering

Next, we download the data from the Google Sheet where it is collected.
Dr. Gilmore has stored his Google account credentials in a special environment file that can be accessed by the R command `Sys.getenv("GMAIL_SURVEY")`.

```{r, message=FALSE, echo=TRUE, eval=FALSE}
if (!dir.exists('csv')) {
  message("Creating missing `csv/`.")
  dir.create("csv")
}

options(gargle_oauth_email = Sys.getenv("GMAIL_SURVEY"))
googledrive::drive_auth()

googledrive::drive_download("Openness and Trust in Science (Responses)", path = "csv/survey-01-openness-trust.csv", type = "csv", overwrite = TRUE)
```

The data file has been saved as a comma-separated value (CSV) format data file in a special directory called `csv/`.

<!-- Note that there are five types of callouts, including note, tip, warning, caution, and important. -->
::: {.callout-note}
Because these data *might* contain sensitive or identifiable information, we only keep a local copy and do not share it publicly via GitHub.
This is achieved by adding the name of the data directory to a special `.gitignore` file.
:::

## Cleaning

Next we load the data file and clean it.

```{r, message=FALSE, echo=TRUE}
survey_01 <-
  readr::read_csv("csv/survey-01-openness-trust.csv", show_col_types = FALSE)

# Google Forms puts the full question in the top row of the data file.
# We use the names() function to extract and print the original questions.
survey_01_qs <- names(survey_01)
survey_01_qs
```

For plotting and analyses, it's usually easier to shorten the questions by creating a short name that reflects the underlying idea or construct.
We'll use the `rename()` function from the `dplyr` package for this. 

We first rename the variables from the "Openness to Revising One’s Viewpoint" [subscale](https://seaver.pepperdine.edu/social-science/content/comprehensive-intellectual-humility.pdf) from the Comprehensive Intellectual Humility Scale [@Krumrei-Mancuso2016-qm].

```{r}
survey_01_clean <- survey_01 |>
  dplyr::rename(
    timestamp = "Timestamp",
    changed_when_shown_wrong = "I have at times changed opinions that were important to me, when someone showed me I was wrong.",
    would_change_good_reason = "I am willing to change my position on an important issue in the face of good reasons." ,
    would_changed_new_info = "I am open to revising my important beliefs in the face of new information.",
    would_change_compelling_reason = "I am willing to change my opinions on the basis of compelling reason.",
    would_change_mind_when_made_up = "I’m willing to change my mind once it’s made up about an important topic.",
    comments = "Comments"
  )
```

Now, we rename the variables from the [@Nadelson2014-ym] trust in science and scientists survey.

```{r}
survey_01_clean <- survey_01_clean |>
  dplyr::rename(
    ignore_contradictory_evidence = "Scientists ignore evidence that contradicts their work.",
    theories_are_weak = "Scientific theories are weak explanations.",
    keep_work_secret = "Scientists intentionally keep their work secret.",
    dont_value_others_ideas = "Scientists don't value the ideas of others.",
    dont_care_laypeople_understand = "Scientists don't care if laypersons understand their work.",
    should_trust_work = "We should trust the work of scientists.",
    should_trust_honesty = "We should trust that scientists are being honest in their work.",
    should_trust_ethical = "We should trust that scientists are being ethical in their work.",
    more_understanding_more_trust = "People who understand science more have more trust in science.",
    trust_explain_natural_world = "We can trust science to find the answers that explain the natural world.",
    cant_trust_biased = "We cannot trust scientists because they are biased in their perspectives.",
    protect_each_other_when_wrong = "Scientists will protect each other even when they are wrong.",
    wont_consider_contradictory_ideas = "We cannot trust scientists to consider ideas that contradict their own.",
    sacrifice_others_to_advance = "Today's scientists will sacrifice the well being of others to advance their research.",
    cant_trust_moves_slowly = "We cannot trust science because it moves too slowly.",
    change_minds_undermines_trust = "When scientists change their mind about a scientific idea it diminishes my trust in their work.",
    share_findings_dont_like = "We can trust scientists to share their discoveries even if they don't like their findings.",
    make_life_better = "I trust that the work of scientists is to make life better for people.",
    theories_trustworthy = "Scientific theories are trustworthy.",
    hypotheses_just_guesses = "When scientists form a hypothesis they are just guessing.",
    trust_find_tech_solutions = "I trust scientists can find solutions to our major technological problems."
  )
```

Now, let's look at the names to confirm they all got changed.

```{r}
names(survey_01_clean)
```

### Data dictionary

We'll pause here to start building a [data dictionary](https://en.wikipedia.org/wiki/Data_dictionary), a file that explains the origin, format, and usage of our dataset.

```{r}
survey_01_data_dictionary <- tibble::tibble(question = survey_01_qs, short_name = names(survey_01_clean),
                                            reference = c(NA, rep("krumrei-mancuso-2016", 5), rep("nadelson-2014", 21), NA))
```

We'll add other items to the data dictionary later.

### Filtering out irrelevant responses

We should omit the first response in the dataset.
That was the one Dr. Gilmore used to generate a Google Sheet, and isn't real data.

```{r}
n_responses <- dim(survey_01_clean)[1]

if (n_responses > 1) {
  survey_01_clean <- survey_01_clean[2:n_responses]  
} else {
  message("No 'non-test' responses yet. Leaving data file unchanged.")
}
```

