---
title: "2023-10-06 Supplement"
author: "Rick Gilmore"
bibliography: [../include/bib/packages.bib, ../include/bib/psu-repro.bib]
csl: ../include/bib/apa.csl
---

## Purpose

This document simulates some data to illustrate Study 1 from [@simmons_false-positive_2011].

## Generate data

There are two song groups:

```{r}
song <- c("kid", "ctrl")
song <- rep(song, each = 15)
song
```

One covariate, father's age.
Let's say that these are uniformly distributed between 40 and 55.

```{r}
father_age <- runif(30, 40, 55)
hist(father_age)
```

And one focus measure, “How old do you feel right now?”, with five options: very young, young, neither young nor old, old, and very old.
Let's give these numbers between 1 and 5, and let's say that the average rating for the "kid" group was 3.75 and the average rating for the "ctrl" group was 3.
We'll assume that these are normally distributed.

```{r}
old_feeling <- c(rnorm(15, 3.75), rnorm(15, 3))
hist(old_feeling)
```

## Visualize data

We'll make a data frame.

```{r}
false_pos_study_1 <- data.frame(sub_id = 1:30, song, father_age, old_feeling)
str(false_pos_study_1)
```
Now, let's plot this to show how analysis of covariance (ANCOVA) might work.

```{r fig-ancova, fig.cap="Illustration of a simulated ANCOVA based on [@simmons_false-positive_2011]"}

library(tidyverse)

false_pos_study_1 |>
  ggplot() +
  aes(x = father_age, y = old_feeling, color = song) +
  geom_point() +
  geom_smooth(method = lm) # linear fit
```

```{r}
glm_ancova <- aov(formula = old_feeling ~ song + father_age, data = false_pos_study_1)
summary(glm_ancova)
```
Now, let's imagine that `father_age` wasn't entirely random for some reason.

```{r}
father_age <- (69:40) + rnorm(30, 0, 1)
plot(1:30, father_age)
hist(father_age)
```

```{r}
false_pos_study_1a <- data.frame(sub_id = 1:30, song, father_age, old_feeling)

false_pos_study_1a |>
  ggplot() +
  aes(x = father_age, y = old_feeling, color = song) +
  geom_point() +
  geom_smooth(method = lm) # linear fit

glm_ancova <- aov(formula = old_feeling ~ song + father_age, data = false_pos_study_1a)
summary(glm_ancova)
```

