---
title: "Survey 03"
params:
  update_data: TRUE
---

## Purpose

This page documents the data processing steps involved with Survey 03 in PSYCH 490.009.
The survey questions were adapted from those discussed in [@Cruwell2019-nz].

The page also serves as a learning opportunity for exploring how to generate a set of plots using functional programming techniques.

## Survey

Link: <https://forms.gle/PFTGCMwmZZj1TtJS7>

<iframe src="https://docs.google.com/forms/d/e/1FAIpQLSf9w7CHjZD7jon1-_BsR_ZZHlIEVdKY0x4nJwh7v8AyXlpUmg/viewform?embedded=true" width="640" height="3972" frameborder="0" marginheight="0" marginwidth="0">Loading…</iframe>

## Preparation

First, we load the external packages (groups of R commands) that we will be using.

```{r include=FALSE}
library('tidyverse')
library('ggplot2')
library('dplyr')
library('tidyr')
library('stringr')
```

## Gathering

Next, we download the data from the Google Sheet where it is collected.
Dr. Gilmore has stored his Google account credentials in a special environment file that can be accessed by the R command `Sys.getenv("GMAIL_SURVEY")`.

```{r, message=FALSE, echo=TRUE}
if (!dir.exists('csv')) {
  message("Creating missing `csv/`.")
  dir.create("csv")
}

if (params$update_data) {
  options(gargle_oauth_email = Sys.getenv("GMAIL_SURVEY"))
  googledrive::drive_auth()
  
  googledrive::drive_download(
    "PSYCH 490.000 Fall 2023 Survey 03 (Responses)",
    path = "csv/survey-03-crewell-et-al.csv",
    type = "csv",
    overwrite = TRUE
  )
  
  message("Data updated.")
} else {
  message("Using stored data.")
}
```

The data file has been saved as a comma-separated value (CSV) format data file in a special directory called `csv/`.

<!-- Note that there are five types of callouts, including note, tip, warning, caution, and important. -->

::: {.callout-note}

Because these data *might* contain sensitive or identifiable information, we only keep a local copy and do not share it publicly via GitHub.
This is achieved by adding the name of the data directory to a special `.gitignore` file.

:::

## Cleaning

Next we load the data file and clean it.

```{r, message=FALSE, echo=TRUE}
survey_03 <-
  readr::read_csv("csv/survey-03-crewell-et-al.csv", show_col_types = FALSE)

# Google Forms puts the full question in the top row of the data file.
# We use the names() function to extract and print the original questions.
survey_03_qs <- names(survey_03)
survey_03_qs
```

### Clean/shorten names

For plotting and analyses, it's usually easier to shorten the questions by creating a short name that reflects the underlying idea or construct.
We'll use the `rename()` function from the `dplyr` package for this. 

```{r}
new_names <-
  c(
    "timestamp",
    "psych_problems_replicating",
    "replication_problem_psych_only",
    "incentives_undermine",
    "low_power_incorrect",
    "replicators_unqualified",
    "report_all_measures_important",
    "decide_n_before_important",
    "share_data_important",
    "collection_analysis_decisions_affect",
    "report_null_findings_important",
    "counterintuitive_results_important"
  )

# Swap out old (long) names for new (short) names
long_names <- names(survey_03)
names(survey_03) <- new_names
```

### Data dictionary

We'll pause here to start building a [data dictionary](https://en.wikipedia.org/wiki/Data_dictionary), a file that explains the origin, format, and usage of our dataset.

```{r}
# Make new data frame with long and short names for reference
survey_03_data_dictionary <-
  tibble::tibble(q_long = long_names, q_short = new_names)

survey_03_data_dictionary |>
  knitr::kable(format = 'html') |>
  kableExtra::kable_classic()
```

We'll add other items to the data dictionary later.

## Visualizations

```{r}
# Retrieve the "long" question from the survey_03 data dictionary
retrieve_long_q <- function(this_q_short, data_dict = survey_03_data_dictionary) {
  assertthat::is.string(this_q_short)
  data_dict |>
    filter(q_short == this_q_short) |>
    select(q_long) |>
    as.character()
}

retrieve_long_q("psych_problems_replicating")
```

```{r}
my_hist <- function(data, var, q_long = "test") {
  data |>
    ggplot() +
    aes(x = {{var}}) +
    geom_histogram() +
    xlim(c(.5,5.5)) +
    ggtitle(q_long) +
    xlab("Rating")
}

my_hist_q <- function(var, data) {
  this_q <- retrieve_long_q(var)
  my_hist(data, .data[[var]], this_q)
}

my_hist_q(var = "replication_problem_psych_only", data = survey_03)
```

Now, we're ready to print histograms for all of the data.

```{r}
my_vars <- names(survey_03)[2:dim(survey_03)[2]]

purrr::map(my_vars, my_hist_q, survey_03)
```

::: {.callout-note "To do"}

Some titles are too long to fit over the plots.
This should be fixed in a future version.

As one approach, I need a function that splits a long string by inserting a line feed `\n` character at the break point(s).

Another approach would be to generate Rmarkdown code that includes the long question as a figure caption (`fig.cap="My title"`).

:::

